---
title: "My First Report"
author: "Jade Lee"
format:
  html:
    embed-resources: true
---

## Welcome to your practice report

This is a **Quarto** document. It mixes ordinary writing with live code, so your
report and your analysis always stay in sync. When you click **Render**, Quarto
runs the code below and builds a finished HTML page.

Try it now: change `author: "Your Name"` in the header above to your own name,
then click **Render**.

## A first look at the data

The code chunk below reads a small practice dataset from the `data/` folder and
shows the first few rows.

```{r}
patients <- read.csv("data/patients.csv")
head(patients)
```

## A first chart

```{r}
hist(patients$age,
     main = "Age distribution of practice patients",
     xlab = "Age (years)",
     col  = "steelblue")
```

## Your turn

```{r}
summary(patients)
```


```{r}
# My added R code: show only patients with diabetes
patients[patients$diabetes == "yes", ]
```

The histogram shows that most patients are between 40 and 70 years old, with only 1 patient over 70. 
Based on the summary and the table of patients with diabetes that I added above, all 4 patients aged 60 or older in this 12-patient sample have diabetes.
However, with only 12 patients, and synthetic data, this cannot be seen as a meaningful result.

## My Notes

- Quarto does **not** re-render automatically, so I need to render again after every edit.
- With `embed-resources: true`, Quarto packs the chart, styles, and scripts into one HTML file, which is why my first commit added over 2,600 lines.
- I used Claude (AI) to help me write the R code for the diabetes table and to double-check that my notes match the lecture, and I verified the table by comparing it with `data/patients.csv`.
