---
title: "Assignment 2 Starter Analysis"
format: html
---

## Setup

```{r}
#| message: false
#| warning: false
library(tidyverse)  # AI-EDIT(2026-06-23): tidyverse-default — consolidated core library() calls into library(tidyverse)
```

## Load Data

```{r}
nhanes <- read_csv("examples/nhanes-equity/data/nhanes_equity_v6.csv", show_col_types = FALSE)
```

## Inspect

```{r}
dim(nhanes)
names(nhanes)
glimpse(nhanes)
colSums(is.na(nhanes))
```

- The dataset has 105,626 rows (people) and 18 columns (variables).

- Seven columns have no missing values: Cycle, WeightMEC, Strata, PSU, Gender, Race, and Age (by colSums(is.na(nhanes))).

- Education and Marital are missing for almost half of the people.
so I will not simply drop these rows. If I use these columns,
I will first analyze who is missing, keep the missing values as a separate group,
and then decide how to handle them.

- BMI, Height, PIR, and IncomeGroup are missing for fewer than 1 in 10 people.
For these columns, I will skip missing values when calculating (na.rm = TRUE)
and show the number of missing values and the number of people (n) in each group.


## Summarize

Replace this example with your own summary if you choose a different question.

```{r}
bmi_summary <- nhanes |>
  filter(!is.na(BMI), !is.na(IncomeGroup)) |>
  group_by(IncomeGroup) |>
  summarise(
    n = n(),
    mean_bmi = mean(BMI),
    .groups = "drop"
  )

bmi_summary
```

## Function

```{r}
mean_with_n <- function(x) {
  tibble(
    n = sum(!is.na(x)),
    mean = mean(x, na.rm = TRUE)
  )
}

mean_with_n(nhanes$BMI)
```

## Save Output

```{r}
dir.create("assignments/assignment02-r-with-ai/submission/output", recursive = TRUE, showWarnings = FALSE)
write_csv(bmi_summary, "assignments/assignment02-r-with-ai/submission/output/summary_table.csv")
```

## AI Audit Note

Add comments explaining which parts were drafted or debugged with AI and how you verified the result.
