---
title: "Assignment 2 Adult BMI Across NHANES Survey Cycles"
format: html
---

## Setup

```{r}
#| message: false
#| warning: false
library(tidyverse)  # AI-EDIT(2026-06-23): tidyverse-default — consolidated core library() calls into library(tidyverse)
```

## Load Data

```{r}
nhanes <- read_csv("examples/nhanes-equity/data/nhanes_equity_v6.csv", show_col_types = FALSE)
```

## Inspect

```{r}
dim(nhanes)
names(nhanes)
glimpse(nhanes)
colSums(is.na(nhanes))
```

- The dataset has 105,626 rows (people) and 18 columns (variables).

- Seven columns have no missing values: Cycle, WeightMEC, Strata, PSU, Gender, Race, and Age (by `colSums(is.na(nhanes))`).

- Education and Marital are missing for almost half of the people.
So I will not simply drop these rows. If I use these columns,
I will first analyze who is missing, keep the missing values as a separate group,
and then decide how to handle them.

- BMI, Height, PIR, and IncomeGroup are missing for fewer than 1 in 10 people.
For these columns, I will skip missing values when calculating (na.rm = TRUE)
and show the number of missing values and the number of people (n) in each group.


## Summarize

Among adults (age 20 and over), did mean BMI increase across survey cycles?

```{r}
# AI prompt: "Among adults 20+, mean BMI by Cycle with n and missing count, plus change from the first cycle"
# Verified: column names match names(nhanes); n across cycles adds up to 58,462,
#           which equals nhanes |> filter(Age >= 20) |> nrow(); na.rm = TRUE is used
# AI prompt: "summary_table.csv shows -0.1999999999999993. Fix it in the code, not the file"
# Verified: after rounding, summary_table.csv shows -0.2, 0.4, ... and matches bmi_summary

bmi_summary <- nhanes |>
  filter(Age >= 20) |>
  group_by(Cycle) |>
  summarise(
    n = n(),
    n_missing_bmi = sum(is.na(BMI)),
    mean_bmi = mean(BMI, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    change_from_first = round(mean_bmi - first(mean_bmi), 1),
    mean_bmi = round(mean_bmi, 1)
  )

bmi_summary
```

Among adults, mean BMI rose from 28.4 in 1999-2000 to 29.8 in 2021-2023,
an increase of about 1.5.
The increase was not steady: mean BMI dipped slightly in 2001-2002 and 2011-2012,
and it was highest in 2017-2018 (29.9).
Few adults were missing BMI in each cycle, except 2001-2002, which had more missing values.
There is no 2019-2020 cycle in this dataset, so the last gap is longer.
These are unweighted sample means, so they describe this sample, not all U.S. adults.

## Function

```{r}
# AI prompt: "Turn my adult BMI-by-cycle summary into a function that works for any numeric column"
# Verified: summarise_by_cycle(nhanes, "BMI") gives the same n, missing counts,
#           and means as bmi_summary in the Summarize section
# AI prompt (GitHub Copilot Chat): "Explain what this function does, line by line. Keep it short."
# Verified: compared each line of the explanation with the code, and checked it against
#           the BMI output table (n, n_missing, mean_value)

summarise_by_cycle <- function(data, column_name) {
  data |>
    filter(Age >= 20) |>
    group_by(Cycle) |>
    summarise(
      n = n(),
      n_missing = sum(is.na(.data[[column_name]])),
      mean_value = mean(.data[[column_name]], na.rm = TRUE),
      .groups = "drop"
    )
}

summarise_by_cycle(nhanes, "BMI")
summarise_by_cycle(nhanes, "Waist")
summarise_by_cycle(nhanes, "WHtR")
```

I turned the Summarize code into a function, summarise_by_cycle(),
so I can repeat the same adult-by-cycle summary for any numeric column.
For BMI, it gives the same results as the table above.
Waist and WHtR show the same pattern as BMI:
mean waist rose from about 96.7 cm in 1999-2000 to about 101 cm in 2021-2023,
and mean WHtR rose from 0.582 to 0.606.

## Save Output

```{r}
dir.create("assignments/assignment02-r-with-ai/submission/output", recursive = TRUE, showWarnings = FALSE)
write_csv(bmi_summary, "assignments/assignment02-r-with-ai/submission/output/summary_table.csv")
```


## AI Audit Note

I used Claude and GitHub Copilot Chat to explain the assignment and code, and to draft code.
I chose my own question, whether adult mean BMI increased across survey cycles,
and asked Claude to design the code for it.
I checked that the column names exist in names(nhanes) and that n across all cycles
adds up to 58,462, the number of adults (Age >= 20).
When the saved CSV showed long decimals such as -0.1999999999999993,
Claude suggested rounding change_from_first, so I fixed the code instead of the CSV file.
Claude then turned this code into the function summarise_by_cycle(),
and I checked that it gives the same n, missing counts, and mean BMI as bmi_summary.
Finally, I asked GitHub Copilot Chat to explain the function line by line
and compared its explanation with the code and with the BMI output table.